Parsing Postfix Log Files

### 2009-04-27 doc/logparser-results.tex
Changes made to chapter 6 after proofreading.

Efficiency is an obvious concern when the parser routinely needs to parse
large log files.  The mail server that generated the \numberOFlogFILES{}
log files used in this chapter is a production mail server handling mail
for a university department; the benefit of using this mail server is that
its log files exhibit the idiosyncrasies and peculiarities that a mail
server in the wild must deal with, but the downside is that significantly
altering the configuration to accommodate this project is not an option.
\Graphref{Mails received per day} and \tableref{Number of mails received
per day: statistics} show that, as expected, it accepts far more mails on
weekdays than at weekends.  Note that mail generated on the server (bounce
notifications, mail re-injected for forwarding, mail sent from mailing
lists) does not contribute to these figures: in particular, the vast amount
of mail resulting from the mail loops noticeable in later graphs is not
reflected in these figures.

The standard deviation for mails received on weekend days is
quite high, because approximately four times the usual number of mails was
received on the weekend contained in the $16^{th}$ and $17^{th}$ log files.
The standard deviation for all days is high because the high number of
mails received during log files 16, 17, 35, 36, 39, and 40.  Interestingly,
the extra mail received during log files 35, 36, 39, and 40 stopped during
log files 37 and 38, which may indicate that it was a virus infected office
machine that was turned off over the weekend.  Unfortunately, the source
cannot be investigated properly because all the mails in question were
relayed via the department's secondary MX (a backup mail server, hosted
elsewhere), so the log files showing the original source were unavailable.

@@

\Graphref{parsing time vs log file size vs number of log lines factor}
plots both the ratio of log file size to parsing time, and the ratio of
number of log lines to parsing time (higher is more efficient in both
cases); \tableref{parsing time vs log file size vs number of log lines
factor table} shows the same ratios for different groups of log files.  The
ratios are quite tightly banded, except they increase (i.e.\ improve) for
log files 22 \& 62--68, despite their larger than usual size.  Both groups
of log files are much larger than usual because of mail loops caused by
users who set up mail forwarding incorrectly, resulting in a very different
distribution of log lines: normally most log lines are generated by mail
delivery attempts from other hosts, but when the mail loops occurred most
of the log lines resulted from failed delivery of mail re-injected for
forwarding.  The Postfix components that generated most of the log lines
during the mail loops have fewer associated rules than the Postfix
components whose log lines normally make up the bulk of each log file, so
the mean number of rules used per log line is lower for these log files, as
is the mean parsing time per log line.  \Tableref{Number of rules per
Postfix component} shows the number of rules for each Postfix component;
\graphref{Mean number of rules used per log line} shows the drop in the
mean number of rules used per log line for log files containing a mail
loop, and \graphref{Mean number of rules used per log line for each Postfix
component} shows the mean number of rules used per log line for each
Postfix component for all log files.

The three plots on \graphref{parsing time vs log file size vs number of log
lines graph} are quite close together, and the graph can be difficult to
read.  Log files 62--68 show a large gap between log file size and number
of log lines, whereas those plots appear to have a small gap or to overlap
for all the other log files, suggesting that the ratio of log file size to
number of log lines is lower for log files 62--68.  This appearance is
misleading: \graphref{Mean log line size in bytes} shows the mean log line
size for all \numberOFlogFILES{} log files, and it reduces for log files 22
\& 62--68, i.e.\ there are more log lines per MB in those log files.

@@

\parsernames{} ruleset has \numberOFrules{} rules: the top 10\% recognise
\input{build/include-top-ten-hits}\% of the log lines in the
\numberOFlogFILES{} log files, with the remaining log lines split across
the other 90\% of the rules, as shown in \graphref{rule hits graph}.
Assuming that the distribution of log lines is reasonably consistent across
log files, \parsernames{} efficiency should benefit from using rules that
recognise log lines more frequently before it uses rules that recognise log
lines less frequently.  To test this hypothesis, three full test runs were
performed with different rule orderings:

@@

The parsing times of the three orderings are plotted against log file size
in \graphref{Parsing time plotted against log file size for optimal,
shuffled, and reverse orderings}; \graphref{Parsing time of optimal and
reverse orderings relative to shuffled ordering graph} shows the parsing
times of optimal and reverse orderings relative to shuffled ordering, and
mean relative parsing times for different groupings of log files are given
in \tableref{Parsing time of optimal and reverse orderings relative to
shuffled ordering table}.  This optimisation provides a mean reduction in
parsing time of
\input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
with normal log files.  \Tableref{Parsing time of optimal and reverse
orderings relative to shuffled ordering table} shows that differences in
rule ordering have less effect on parsing time when parsing log files 22 \&
62--68, because of the different distribution of log lines in those log
files.  A careful examination of \graphref{Parsing time of optimal and
reverse orderings relative to shuffled ordering graph} shows that, for the
first log file only, optimal and reverse orderings perform identically:
this is because the hits field of each rule is zero for the first log file,
so optimal and reverse orderings produce identical rule orderings.  For the
first log file, shuffled ordering is the most efficient of the three, but
that is accidental and cannot be relied upon.

@@

@@

Once again, the difference between the oracle and optimal ordering is at
its lowest when parsing log files resulting from a mail loop (log files 22
\& 62--68), because the mean number of rules used per log line is lower
(see \graphref{Mean number of rules used per log line}).  The performance
of optimal ordering relative to the oracle is much worse when parsing the
first log file than for the remainder of the log files, because the hits
field of every rule starts at zero, so optimal ordering does not provide
any benefit for the first log file; the oracle, in contrast, is flawless
for every log file.

@@

Clearly, the increased number of rules causes a noticeable performance
decrease with reverse ordering, and a lesser decrease with shuffled
ordering, whereas optimal ordering shows scant change.  Log files 22 \&
62--68 show much smaller increases in parsing time than other log files do,
because most of the log lines in those log files are produced by Postfix
components with few rules, so removing unnecessary rules has little effect
on the total number of rules used; \tableref{Number of rules per Postfix
component in the maximum and minimum rulesets} shows the number of rules
per Postfix component for each ruleset.  Once again, \graphref{Percentage
parsing time increase when using the maximum ruleset instead of the minimum
ruleset} shows that optimal and reverse orderings have identical
performance for the first log file, because the hits field of every rule
starts at zero, and the two rule orderings are identical.

@@

A test run using optimal ordering was performed as described in
\sectionref{parser efficiency}, with one difference: regexes were not
compiled and cached when the ruleset was loaded from the database, so
Perl's regex engine had to compile each regex each time it was used when
trying to recognise a log line.  \Graphref{Percentage increase in parsing
time when not caching compiled regexes graph} shows the effect that not
caching compiled regexes has on parser performance, with mean and standard
deviation in \tableref{Percentage increase in parsing time when not caching
compiled regexes table}.  For typical log files, the mean increase in
parsing time when not caching compiled regexes is
\input{build/include-cached-regexes-vs-discarded-regexes-mean}; looking at
it from the opposite direction, caching compiled regexes reduces parsing
time by
\input{build/include-cached-regexes-vs-discarded-regexes-mean-reduction}.
Caching compiled regexes is probably the single most effective optimisation
possible in \parsername{}, and was quite simple to implement: the framework
compiles each rule's regex when the ruleset is loaded, and uses the
compiled regex instead of the source regex when recognising log lines.  As
seen previously, log files 22 \& 62--68 do not suffer such a large increase
in parsing time when the optimisation is disabled; this is because, on
average, fewer rules are used per log line, so fewer regexes are compiled
per log line for those log files.

The increase in parsing time when parsing the first log file is much
greater than for the other log files (see \graphref{Percentage increase in
parsing time when not caching compiled regexes graph}); again, this is
because every rule's hits field starts at zero, so optimal ordering is less
efficient than usual, the mean number of rules used for each log line is
higher than usual, and more regexes will need to be compiled when
recognising each log line.  Parsing time is plotted against log file size
for cached and discarded regexes in \graphref{Parsing time plotted against
log file size for cached regexes and discarded regexes}; the plot for
discarding compiled regexes is quite uneven compared to caching compiled
regexes --- some log files are particularly slow, whereas others are not as
badly affected, but these differences have not been investigated.

@@

\showgraph{build/graph-file-size-optimal-discarded}{Parsing time plotted
against log file size for cached regexes and discarded regexes}{Parsing
time plotted against log file size for cached regexes and discarded
regexes}

@@

    \item Even with the optimisations described in this chapter enabled,
        recognising log lines still accounts for two thirds of the parser's
        execution time.  \Graphref{Percentage of parsing time spent
        recognising log lines graph} shows the percentage of parsing time
        spent recognising log lines for each of the \numberOFlogFILES{} log
        files, with mean and standard deviation shown in
        \tableref{Percentage of parsing time spent recognising log lines
        table}; for normal log files,
        \input{build/include-percentage-time-spent-recognising-log-lines-mean}
        of parsing time is spent recognising log lines.  To measure how
        long recognition of log lines takes, a full test run was performed
        as previously described in \sectionref{parser efficiency}, but with
        one difference: actions were not invoked when a log line was
        recognised, so the parsing time did not include the time normally
        taken by invocation of actions; the mean parsing times for each log
        file were subtracted from the mean parsing times from a normal test
        run to obtain the parsing time taken by actions.  The parsing time
        exclude the time taken for framework initialisation, loading and
        saving state tables, and loading the ruleset: in ad far as
        possible, the times are just for recognising log lines or
        recognising log lines plus invoking actions.

        Optimal ordering reduces parsing time by
        \input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
        relative to shuffled ordering; actions occupy
        \input{build/include-percentage-time-spent-processing-log-lines-mean}
        of parsing time, less if any optimisations are disabled, so
        optimising actions could not possibly provide as big a performance
        increase as optimally ordering rules.  Similarly, caching compiled
        regexes provides an
        \input{build/include-cached-regexes-vs-discarded-regexes-mean-reduction}
        reduction in parsing time; not invoking actions at all reduces the
        most optimised parsing time by less than half that.  If processing
        of recognised log lines was optimised to 1\% of its original
        parsing time, it would be only slightly more effective than optimal
        ordering, but it would be vastly more difficult to implement;
        optimising actions would not reduce parsing time enough to justify
        the amount of effort required.

@@

Each Postfix component whose log lines are of interest must have at least
one rule that recognises its log lines, or all of its log lines will be
silently skipped; in the extreme case of an empty ruleset the parser would
skip every log line.  \parsername{} skips log lines from programs that do
not have any associated rules, because there may be any number of log lines
from other programs in the log file, and some Postfix components do not
produce any log lines of interest.  \parsername{} does not parse log lines
from non-Postfix programs, e.g.\ Amavisd-new or SpamAssassin; it could
easily be extended to do so, if a method could be developed to correctly
associate such log lines with existing state table entries.  To correctly
recognise all log lines, there must be a rule to recognise each log line
variant produced by each Postfix component; if a log line is not recognised
the parser will issue a warning, to inform the user that they need to
extend their ruleset.  Each rule's regex should be as specific and precise
as possible, to ensure accurate parsing: a rule with a regex that matches
zero or more of any character will recognise every log line, but not in a
meaningful way; most log lines contain fixed strings, so this is not a
problem in practice.

@@

The proportion of mail delivery attempts that are correctly understood and
reconstructed is much more difficult to determine accurately than the
proportion of log lines that are correctly recognised.  The parser can dump
its state tables in a human readable form; examining those tables with
reference to the log files is the best way to detect mails that were not
reconstructed properly; many of the complications discussed in
\sectionref{complications} were detected in this way.  \parsername{} issues
warnings when it detects any errors or discrepancies, e.g.\ when a queueid
is reused but the previous mail remains in the state tables, when a queueid
or \acronym{pid} is not found in the state tables, or when an entry in the
state tables does not include sufficient data to satisfy the database
schema.  The parser should produce few or no warnings during parsing, and
when finished parsing the state tables should only contain entries for
mails that have log lines in subsequent log files.  There will often be
warnings about a missing queueid or \acronym{pid} when parsing the first
few thousand log lines, because some earlier log lines for those connections
or mails are in a previous log file; loading saved state tables from the
previous log file will solve this problem.

@@

    \item 9 mails whose initial log lines were not present in the log
        files.  Some of those mails did not produce warnings because they
        resemble child mails waiting to be tracked with a parent; see
        \sectionref{Re-injected mails} for details.

@@

    \item \parsername{} performs many checks to detect known problems,
        e.g.\ if a mail is missing required data.  No such warnings are
        produced during the test runs other than the five described above.

@@

    \item When dealing with the complications described in
        \sectionref{complications}, the solutions are as specific and
        restrictive as possible, with the goal of minimising the number of
        false positives.  In addition, the solution to the complication
        described in \sectionref{out of order log lines} imposes conditions
        that every reconstructed mail must comply with to be acceptable,
        not just the five mails that exhibit that complication.

@@

This chapter evaluated \parsername{} on two criteria: efficiency, and
coverage of Postfix log files.  The former began by describing the mail
server the log files were taken from, the computer used to generate the
statistics in this chapter, and the characteristics of the
\numberOFlogFILES{} log files used to generate statistics and test
coverage, including why performance is better on the larger log files.  The
framework optimises the order in which rules are used when trying to
recognise each log line, and the effect that optimisation has on parsing
time is explored; this is followed by comparing optimal ordering with an
oracle.  How \parsername{} scales as the size of the ruleset increases is
addressed, followed by a description and analysis of the simplest and most
effective of the optimisations, caching compiled regexes, and the
efficiency evaluation concludes with an examination of where parsing time
is spent: recognising log lines or processing them?

>>>>

Efficiency is an obvious concern when the parser routinely needs to parse
large log files.  The mail server that generated the \numberOFlogFILES{}
log files used in this chapter is a production mail server handling mail
for a university department; the benefit of using this mail server is that
its log files exhibit the idiosyncrasies and peculiarities that a mail
server in the wild must deal with, but the downside is that significantly
altering its Postfix configuration to accommodate this project is not an
option.  \Graphref{Mails received per day} and \tableref{Number of mails
received per day: statistics} show that, as expected, it accepts far more
mails on weekdays than at weekends.  Note that mail generated on the server
(e.g.\ bounce notifications, mail re-injected for forwarding, mail sent
from mailing lists) does not contribute to these figures: in particular,
the vast amounts of mail resulting from the mail loops noticeable in later
graphs are not reflected in these figures.

The standard deviation for mails received on weekend days is quite high,
because approximately four times the usual number of mails were received on
the weekend contained in log files 16 \& 17.  The standard deviation for
all days is high because of the unusually high number of mails received
during log files 16, 17, 35, 36, 39 \& 40.  Interestingly, the extra mails
received during log files 35, 36, 39 \& 40 stopped during log files 37 \&
38, which may indicate that the source was a virus infected office machine
that was turned off over the weekend.  Unfortunately, the source cannot be
investigated properly because all of the mails in question were relayed via
the department's secondary \acronym{MTA} (a backup mail server, hosted
elsewhere), so the log files showing the original source were unavailable.
Relaying spam mail through a secondary \acronym{MTA} is a common practice
amongst spam senders, because historically, primary \acronyms{MTA} had
better anti-spam defences than secondary \acronyms{MTA}.

@@

\Graphref{parsing time vs log file size vs number of log lines factor}
plots both the ratio of log file size to parsing time, and the ratio of
number of log lines to parsing time (higher is more efficient in both
cases); \tableref{parsing time vs log file size vs number of log lines
factor table} shows mean and standard deviation.  The ratios are quite
tightly banded, except they increase (i.e.\ improve) for log files 22 \&
62--68, despite their larger than usual size.  The log files in both groups
are much larger than usual because of mail loops caused by users who set up
mail forwarding incorrectly, resulting in a very different distribution of
log lines: normally most log lines are generated by mail delivery attempts
from other hosts, but during the mail loops most of the log lines resulted
from failed delivery of bounce notifications re-injected for forwarding.
The Postfix components that generated most of the log lines during the mail
loops have fewer associated rules than the Postfix components whose log
lines normally make up the bulk of each log file, so the mean number of
rules used per log line is lower for these log files, as is the mean
parsing time per log line.  \Tableref{Number of rules per Postfix
component} shows the number of rules for each Postfix component;
\graphref{Mean number of rules used per log line} shows the drop in the
mean number of rules used per log line for log files containing a mail
loop, and \graphref{Mean number of rules used per log line for each Postfix
component} shows the mean number of rules used per log line for each
Postfix component for \numberOFlogFILES{} log files.

The three plots on \graphref{parsing time vs log file size vs number of log
lines graph} are quite close together, and the graph can be difficult to
read.  Log files 62--68 show a large gap between log file size and number
of log lines, whereas those plots appear to have a small gap for all the
other log files, suggesting that the ratio of number of log lines to log
file size is lower for log files 62--68 (i.e.\ fewer log lines per MB).
This appearance is misleading: \graphref{Mean log line size in bytes} shows
the mean log line size for all \numberOFlogFILES{} log files, and it
reduces for log files 22 \& 62--68, i.e.\ there are more log lines per MB
in those log files.

@@

\parsernames{} ruleset has \numberOFrules{} rules: the top 10 rules
recognise \input{build/include-top-ten-hits}\% of the log lines in the
\numberOFlogFILES{} log files, with the remaining log lines split across
the other \numberOFrulesMINUSten{} of the rules, as shown in \graphref{rule
hits graph}.  Assuming that the distribution of log lines is reasonably
consistent across log files, \parsernames{} efficiency should benefit from
using rules that recognise log lines more frequently before it uses rules
that recognise log lines less frequently.  To test this hypothesis, three
full test runs were performed with different rule orderings:

@@

The parsing times of the three orderings are plotted against log file size
in \graphref{Parsing time plotted against log file size for optimal,
shuffled, and reverse orderings}; \graphref{Parsing time of optimal and
reverse orderings relative to shuffled ordering graph} shows the parsing
times of optimal and reverse orderings relative to shuffled ordering, with
mean and standard deviation in \tableref{Parsing time of optimal and
reverse orderings relative to shuffled ordering table}.  \Graphref{Number
of rules used by optimal, shuffled, and reverse orderings} shows the number
of rules used by each ordering while parsing the \numberOFlogFILES{} log
files.  This optimisation provides a mean reduction in parsing time of
\input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
with normal log files, making it a very worthwhile and effective
optimisation.

\Tableref{Parsing time of optimal and reverse orderings relative to
shuffled ordering table} shows that differences in rule ordering have less
effect on parsing time when parsing log files 22 \& 62--68, because of the
different distribution of log lines in those log files.  A careful
examination of \graphref{Parsing time of optimal and reverse orderings
relative to shuffled ordering graph} shows that, for the first log file
only, optimal and reverse orderings perform identically: this is because
the hits field of each rule is zero for the first log file, so optimal and
reverse orderings produce identical rule orderings.  For the first log
file, shuffled ordering is the most efficient of the three, but that is
accidental and cannot be relied upon.

@@

\showgraph{build/graph-number-or-rules-used-by-different-orderings}{Number
of rules used by optimal, shuffled, and reverse orderings}{Number of rules
used by optimal, shuffled, and reverse orderings}

@@

Once again, the difference between the oracle and optimal ordering is at
its lowest when parsing log files resulting from a mail loop (log files 22
\& 62--68), because the mean number of rules used per log line is lower
when parsing these log files (see \graphref{Mean number of rules used per
log line}).  The performance of optimal ordering relative to the oracle is
much worse when parsing the first log file than for the remainder of the
log files, because the hits field of every rule starts at zero, so optimal
ordering does not provide any benefit for the first log file; the oracle,
in contrast, is flawless for every log file.

@@

Clearly, the increased number of rules causes a noticeable performance
decrease with reverse ordering, and a lesser decrease with shuffled
ordering, whereas optimal ordering shows scant change.  Log files 22 \&
62--68 show much smaller increases in parsing time than other log files do,
because most of the log lines in those log files are produced by Postfix
components with few rules, so removing unnecessary rules has little effect
on the total number of rules used; \tableref{Number of rules per Postfix
component in the maximum and minimum rulesets} shows the number of rules
per Postfix component for each ruleset.  Once again, \graphref{Percentage
parsing time increase when using the maximum ruleset instead of the minimum
ruleset} shows that optimal and reverse orderings have identical
performance for the first log file, because the hits field of every rule
starts at zero, and so the two rule orderings are identical for the first
log file.

@@

A test run using optimal ordering was performed as described in
\sectionref{parser efficiency}, with one difference: regexes were not
compiled and cached when the ruleset was loaded from the database, so the
Perl interpreter had to compile each regex each time it was used when
trying to recognise a log line.  \Graphref{Percentage increase in parsing
time when not caching compiled regexes graph} shows the effect that not
caching compiled regexes has on parser performance, with mean and standard
deviation in \tableref{Percentage increase in parsing time when not caching
compiled regexes table}.  For typical log files, the mean increase in
parsing time when not caching compiled regexes is
\input{build/include-cached-regexes-vs-discarded-regexes-mean}; looking at
it from the opposite direction, caching compiled regexes reduces parsing
time by
\input{build/include-cached-regexes-vs-discarded-regexes-mean-reduction}.
Caching compiled regexes is probably the single most effective optimisation
possible in \parsername{}, and was quite simple to implement: the framework
compiles each rule's regex when the ruleset is loaded, and uses the
compiled regex instead of the source regex when recognising log lines.  As
seen previously, log files 22 \& 62--68 do not suffer such a large increase
in parsing time when the optimisation is disabled; this is because, on
average, fewer rules are used per log line, so fewer regexes are compiled
per log line for those log files.

The increase in parsing time when parsing the first log file is much
greater than for the other log files (see \graphref{Percentage increase in
parsing time when not caching compiled regexes graph}); again, this is
because every rule's hits field starts at zero, so optimal ordering is less
efficient than usual, the mean number of rules used for each log line is
higher than usual, and more regexes will need to be compiled when
recognising each log line.  Parsing time is plotted against log file size
for caching and not caching compiled regexes in \graphref{Parsing time
plotted against log file size when caching and not caching compiled
regexes}; the plot for not caching compiled regexes is quite uneven
compared to caching compiled regexes --- some log files are particularly
slow, whereas others are not as badly affected, but these differences have
not been investigated.

@@

\showgraph{build/graph-file-size-optimal-discarded}{Parsing time plotted
against log file size when caching and not caching compiled
regexes}{Parsing time plotted against log file size when caching and not
caching compiled regexes}

@@

    \item Even with the optimisations described in this chapter enabled,
        recognising log lines still accounts for two thirds of the parser's
        execution time.  \Graphref{Percentage of parsing time spent
        recognising log lines graph} shows the percentage of parsing time
        spent recognising log lines for each of the \numberOFlogFILES{} log
        files, with mean and standard deviation shown in
        \tableref{Percentage of parsing time spent recognising log lines
        table}; for normal log files,
        \input{build/include-percentage-time-spent-recognising-log-lines-mean}
        of parsing time is spent recognising log lines.  To measure how
        long recognition of log lines takes, a full test run using optimal
        ordering and caching compiled regexes was performed as previously
        described in \sectionref{parser efficiency}, but with one
        difference: actions were not invoked when a log line was
        recognised, so the parsing time did not include the time normally
        taken by invocation of actions; the mean parsing times for each log
        file were subtracted from the mean parsing times from a normal test
        run to obtain the parsing time taken by actions.  The parsing times
        exclude the time taken for framework initialisation, loading and
        saving state tables, loading the ruleset, and other housekeeping
        tasks: in as far as possible, the times are just for recognising
        log lines or recognising log lines plus invoking actions.

        Optimal ordering reduces parsing time by
        \input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
        relative to shuffled ordering; actions occupy
        \input{build/include-percentage-time-spent-processing-log-lines-mean}
        of parsing time, less if any optimisations are disabled, so
        optimising actions could not possibly provide as big a performance
        increase as optimally ordering rules.  Similarly, caching compiled
        regexes provides an
        \input{build/include-cached-regexes-vs-discarded-regexes-mean-reduction}
        reduction in parsing time; not invoking actions at all reduces the
        most optimised parsing time by less than half that.  Optimising
        actions to 1\% of their original parsing time would be only
        slightly more effective than optimal ordering is, but would be
        vastly more difficult to implement; optimising actions would not
        reduce parsing time enough to justify the amount of effort
        required.

@@

Each Postfix component whose log lines are of interest must have at least
one rule that recognises its log lines, or all of its log lines will be
silently skipped; in the extreme case of an empty ruleset the parser would
skip every log line.  \parsername{} skips log lines from programs that do
not have any associated rules, because there may be any number of log lines
from other programs in the log file, and some Postfix components do not
produce any log lines of interest.  \parsername{} does not parse log lines
from non-Postfix programs, e.g.\ Amavisd-new or SpamAssassin; it could
easily be extended to do so, if a method could be developed to correctly
associate such log lines with existing state table entries.  To correctly
recognise all log lines that are not skipped, there must be a rule to
recognise each log line variant produced by each Postfix component; if a
log line is not recognised the parser will issue a warning, to inform the
user that they need to extend their ruleset.  Each rule's regex should be
as specific and precise as possible, to ensure accurate parsing: a rule
with a regex that matches zero or more of any character will recognise
every log line, but not in a meaningful way; most log lines contain fixed
strings, so this is not a problem in practice.

@@

The proportion of mail delivery attempts that are correctly understood and
reconstructed is more difficult to determine accurately than the proportion
of log lines that are correctly recognised.  The parser can dump its state
tables in a human readable form; examining those tables with reference to
the log files and the database is the best way to detect mails that were
not reconstructed properly; many of the complications documented in
\sectionref{complications} were detected in this way.  \parsername{} issues
warnings when it detects any errors or discrepancies, e.g.\ when a queueid
is reused but the previous mail remains in the state tables, when a queueid
or \acronym{pid} is not found in the state tables, or when an entry in the
state tables does not include sufficient data to satisfy the database
schema.  The parser should produce few or no warnings during parsing, and
when finished parsing the state tables should only contain entries for
mails that have log lines in subsequent log files.  There will often be
warnings about a missing queueid or \acronym{pid} when parsing the first
few thousand log lines, because some earlier log lines for those
connections or mails are in a previous log file; loading state tables saved
when parsing the log file containing those log lines will solve this
problem.

@@

    \item 9 mails whose initial log lines were not present in the log
        files.  Six of those mails did not produce warnings because they
        resemble child mails waiting to be tracked with a parent; see
        \sectionref{Re-injected mails} for details.  Three of these mails
        were missing more log lines, and so did five produce warnings, as
        documented previously.

@@

    \item \parsername{} performs many checks to detect known problems,
        e.g.\ a queueid missing from the state tables.  No such warnings
        are produced during the test runs other than the five described
        previously.

@@

    \item When dealing with the complications described in
        \sectionref{complications}, the solutions are as specific and
        restrictive as possible, with the goal of minimising the number of
        false positives.  In addition, the solution to the complication
        described in \sectionref{out of order log lines} imposes conditions
        that every reconstructed mail must comply with to be acceptable,
        not just the five mails exhibiting that complication.

@@

This chapter evaluated \parsername{} on two criteria: efficiency, and
coverage of Postfix log files.  The former began by describing the mail
server the log files were taken from, the computer used to generate the
statistics in this chapter, and the characteristics of the
\numberOFlogFILES{} log files used to generate statistics and test
coverage, including why performance is better when parsing the larger log
files.  The framework optimises the order in which rules are used when
trying to recognise each log line, and the effect that optimisation has on
parsing time is explored; this is followed by comparing optimal ordering
with an oracle.  How \parsername{} scales as the size of the ruleset
increases is addressed, followed by a description and analysis of the
simplest and most effective of the optimisations, caching compiled regexes,
and the efficiency evaluation concludes with an examination of where
parsing time is spent: recognising log lines or processing them?
